---
title: "Homework 3"
author: "David Edwards"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyr)
library(dplyr)
```


## Problem 3

Redo Problem 4 parts a-d from last time using the tidyverse functions and piping.

a. Sensory data from five operators.    
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat> 
b. Gold Medal performance for Olympic Men's Long Jump, year is coded as 1900=0.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat>  
c. Brain weight (g) and body weight (kg) for 62 species.    
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat>  
d. Triplicate measurements of tomato yield for two varieties of tomatos at three planting densities.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat>  

### Part a

Here is my code for part a.  Didn't really see a need for piping in this one but it was much easier being able to actually read in the data to a table vs parsing it line by line as I had to do the previous week.

```{r}
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
data <- fread(url, fill = TRUE, sep = " ",  skip = 2, col.names = c("Item", "Op1", "Op2",
                                                                    "Op3","Op4","Op5"))
j <-1
for (i in seq(1,dim(data)[1], 3)){
  data[i+1] = c(j, data[i+1,1:5])
  data[i+2] = c(j, data[i+2,1:5])
  j<- j+1
}
data <- gather(data, Operator, Value, -Item)
data
```

### Part b

I don't remember seeing a function in dplyr or tidyr that would have helped with this one (I'm sure I missed something).  However, this time I was able to just read in the file and then combine the columns in a much simpler way than before.

```{r}
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
data <- fread(url, fill = TRUE, skip = 1)
mydf <- data.frame("Year"=c(data$V1,data$V3,data$V5,data$V7[!is.na(data$V7)]),
                   "LongJump"=c(data$V2, data$V4, data$V6, data$V8[!is.na(data$V8)]))
mydf

```

### Part c

Just like part b as this is almost the same problem (in terms of cleaning the data)

```{r}
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
data <- fread(url, fill = TRUE, skip = 1)
mydf <- data.frame("BodyWt"=c(data$V1,data$V3,data$V5[!is.na(data$V5)]),
                   "BrainWt"=c(data$V2, data$V4, data$V6[!is.na(data$V6)]))

mydf
```

### Part d

For this problem I was able to take advantage of the functions in dplyr and tidyr to help tidy up the data set.  The pipes also really helped.  

```{r}
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
data <- fread(url, fill = TRUE, header = T, skip = 1, col.names = c("Treatment", "10000","20000","30000"))
mydf <- bind_rows(data[1,],data[2,])
mydf <- mydf %>% gather(Levels,Values,-Treatment)%>% separate(Values,c("V1","V2","V3"), sep = ",") %>%gather( Version, Values, -(Treatment:Levels)) %>% select(-Version)
mydf

```


## Problem 4

Finish this homework by pushing your changes to your repo.  In general, your workflow for this should be:  

1. In terminal: git pull  -- to make sure you have the most recent local repo  
2. In terminal: do some work  
3. In terminal: git add -- check files you want to commit  
4. In terminal: git commit -- make message INFORMATIVE and USEFUL  
5. In terminal: git push -- this pushes your local changes to the repo  

If you have difficulty with steps 1-5, git is not correctly or completely setup.

**Only submit the .Rmd and .pdf solution files.  Names should be formatted HW3_lastname_firstname.Rmd and HW3_lastname_firstname.pdf**

## Optional preperation for next class:  

TBD -- could be something sent as a class message




