---
title: "STAT 5014 HW 8"
author: "David Edwards"
date: "11/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(dplyr)
library(tidyr)
library(gridExtra)
library(ggplot2)
```

## Problem 3

Read in the data

```{r, eval=FALSE}
#Run all of this just once.
#Read in Data
Ed.data <- read.csv("./Ed_Data.csv", header = T)

#Remove columns that are not needed
Ed.data <- Ed.data[,c(1, 4, 5:69)]

#Determine if any columns are all NA's  
colmunRemove <- c()
for(i in 1:67){
  if(all(is.na(Ed.data[,i]))){
    colmunRemove <- c(colmunRemove, i)
  }
}#Turns out all the columns contain at least 1 value


#Find all rows that contain nothing but NAs
rowRemove <- c()
for(j in 1:nrow(Ed.data)){
  if(all(is.na(Ed.data[j,3:47]))){
    rowRemove <- c(rowRemove, j)
  }
  if(j %% 100 == 0){
    print(j)
  }
}

#Remove the rows
Ed.data <- Ed.data[-rowRemove, ]

#Gather the data
CleanData <- gather(Ed.data, "Year", "Value", -c(ï..Country.Name, Indicator.Code), na.rm = F)

#Again remove rows that contain NAs
CleanData_Rm <- CleanData[!is.na(CleanData$Value),]
save.image(file = "CleanData.RData")
```

```{r}
load(file = "./CleanData.RData")
names(CleanData_Rm) <- c("Country.Name", "Indicator.Code", "Year", "Value")
```

```{r}
#Pick to countries
Iceland <- CleanData_Rm[CleanData_Rm$Country.Name == "Iceland",]
Italy <- CleanData_Rm[CleanData_Rm$Country.Name == "Italy",]

#Combine there data
twoContries <- rbind(Iceland, Italy)
```


```{r}
#Make sure that we compare the same values for each country
twoContries <- spread(twoContries, Indicator.Code, Value)
```

```{r}
#I decided to work with 250 variables to make sure that I have enough data and to show that the ideas work. 
#Since I don't know what any of this data really means I figured which variables I pick doesn't really matter.
twoContries <- twoContries[,1:250]

#Remove the rows that don't contain any information
rowRemove <- c()
for(j in 1:nrow(twoContries)){
  if(all(is.na(twoContries[j,3:250]))){
    rowRemove <- c(rowRemove, j)
  }

}

twoContries <- twoContries[-rowRemove,]
```


```{r}
Iceland <- twoContries[1:9,]
Iceland <- select(Iceland, -c(Country.Name,Year))
Italy <- twoContries[10:18,]
Italy <- select(Italy, -c(Country.Name, Year))

Iceland_Avg <- sapply(Iceland, mean)
Italy_Avg <- sapply(Italy, mean)

two_Avg <- rbind(Iceland_Avg, Italy_Avg)

#two_Avg <- two_Avg[,1:250]

kable(t(two_Avg), "latex", longtable = T, booktabs = T, caption = "Average Values")
```

To answer the rest of the question, there were about 880,000 original data points in the data, after cleaning up there were about 5,000,000 data points.


## Problem 4

This is my solution to the scatter plot with marginal histograms using base R. For this problem I just picked an indicator code and compared the values for the Arab and EU regions. Please ignore the first histogram.  I just created it to get the data out of it to create the next plot but couldn't figure out how to keep it from displaying.   

```{r}
SE.PRM.TENR <- Ed.data[Ed.data$Indicator.Code == "SE.PRM.TENR",]
Arab <- t(SE.PRM.TENR[1,3:47])
EU <- t(SE.PRM.TENR[7,3:47])
H <- hist(EU)
```


```{r}


par(mfrow = c(2,2))
par(mfg=c(1,1))
hist(Arab)
par(mfg=c(2,1))
plot(Arab,EU)

par(mfg=c(2,2))
plot(NULL, type = "n", xlim = c(0, max(H$counts)), ylim = c(range(H$breaks)), xlab = "Frequency", ylab = "EU")
rect(0, H$breaks[1:(length(H$breaks) - 1)], H$counts, H$breaks[2:length(H$breaks)])
```


## Problem 5

Below is my plot using the ggplot2 package.  I found this code on the following site: https://stackoverflow.com/questions/8545035/scatterplot-with-marginal-histograms-in-ggplot2

```{r}
hist_top <- ggplot()+geom_histogram(aes(Arab))
empty <- ggplot()+geom_point(aes(1,1), colour="white")+
         theme(axis.ticks=element_blank(), 
               panel.background=element_blank(), 
               axis.text.x=element_blank(), axis.text.y=element_blank(),           
               axis.title.x=element_blank(), axis.title.y=element_blank())

scatter <- ggplot()+geom_point(aes(Arab, EU))
hist_right <- ggplot()+geom_histogram(aes(EU))+coord_flip()

grid.arrange(hist_top, empty, scatter, hist_right, ncol=2, nrow=2, widths=c(4, 1), heights=c(1, 4))
```

